Outliers and discrepancies

### 2024-10-31 misc/outlier.qmd
fix

@@

# Modify the main table
s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1
  ) %>%
  ungroup() %>%
  filter(color_flag) %>%
  group_by(appliance, family, state, color_flag, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  select(-color_flag) %>%
  # Join with correct values to create is_correct flag
  left_join(correct_lookup, 
            by = c("state", "appliance"),
            suffix = c("", "_correct")) %>%
  mutate(is_correct = abs(as.numeric(family) - family_correct) < 0.1) %>%
  select(-family_correct) %>%
  # Reorganize and format
  select(state, refClass, appliance, family, n, is_correct) %>%  # Swapped order here
  arrange(state, appliance, refClass, family) %>%
  flextable() %>%
  # Merge cells
  merge_v(j = "state") %>%
  merge_v(j = c("appliance", "refClass")) %>%
  # Conditional formatting for both family and count
  color(i = ~ is_correct == TRUE, j = c("family", "n"), color = "darkgreen") %>%
  color(i = ~ is_correct == FALSE, j = c("family", "n"), color = "darkred") %>%
  bg(i = ~ is_correct == TRUE, j = c("family", "n"), bg = "#e6ffe6") %>%
  bg(i = ~ is_correct == FALSE, j = c("family", "n"), bg = "#ffe6e6") %>%
  # Style improvements
  theme_vanilla() %>%
  align(j = c("family", "n"), align = "right") %>%
  border_outer() %>%
  border_inner() %>%
  fontsize(size = 11) %>%
  padding(padding = 4) %>%
  # Add section borders
  hline(i = ~ state != lag(state, default = first(state)), 
        border = officer::fp_border(color = "gray40", width = 2)) %>%
  colformat_double(j = "family", digits = 1) %>%
  set_header_labels(
    state = "State",
    appliance = "Appliance",
    family = "Family",
    refClass = "Reference Class",
    n = "Count",
    is_correct = "Correct Value"
  ) %>%
  autofit()

@@

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(distinct_state_avg = n_distinct(state_avg)) %>%
  ungroup() %>%
  filter(distinct_state_avg > 1) %>%
  group_by(appliance, state_avg, state, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  # Identify outliers (rows where n is 1)
  mutate(is_outlier = n == 1) %>%
  # Reorganize and format
  select(state, refClass, appliance, state_avg, n) %>%
  arrange(state, appliance, refClass, state_avg) %>%
  flextable() %>%
  # Merge cells
  merge_v(j = "state") %>%
  merge_v(j = c("appliance", "refClass")) %>%
  # Conditional formatting for outliers
  color(i = ~ n == 1, j = c("state_avg", "n"), color = "darkred") %>%
  color(i = ~ n > 1, j = c("state_avg", "n"), color = "darkgreen") %>%
  bg(i = ~ n == 1, j = c("state_avg", "n"), bg = "#ffe6e6") %>%
  bg(i = ~ n > 1, j = c("state_avg", "n"), bg = "#e6ffe6") %>%
  # Style improvements
  theme_vanilla() %>%
  align(j = c("state_avg", "n"), align = "right") %>%
  border_outer() %>%
  border_inner() %>%
  fontsize(size = 11) %>%
  padding(padding = 4) %>%
  # Add section borders
  hline(i = ~ state != lag(state, default = first(state)), 
        border = officer::fp_border(color = "gray40", width = 2)) %>%
  colformat_double(j = "state_avg", digits = 0) %>%
  set_header_labels(
    state = "State",
    appliance = "Appliance",
    state_avg = "State Average",
    refClass = "Reference Class",
    n = "Count"
  ) %>%
  autofit()
```

>>>>

## Inspects counts for each distinct value of prior year family use, and state average

@@

# Modify the main table
s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1
  ) %>%
  ungroup() %>%
  filter(color_flag) %>%
  group_by(appliance, family, state, color_flag, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  select(-color_flag) %>%
  # Join with correct values to create is_correct flag
  left_join(correct_lookup, 
            by = c("state", "appliance"),
            suffix = c("", "_correct")) %>%
  mutate(is_correct = abs(as.numeric(family) - family_correct) < 0.1) %>%
  select(-family_correct) %>%
  # Reorganize and format
  select(state, refClass, appliance, family, n, is_correct) %>%  # Swapped order here
  arrange(state, appliance, refClass, family) %>%
  flextable() %>%
  add_header_lines(values = "Family prior year use per utility counts (green=most frequent)") %>%
  bold(part = "header", i = 1) %>%
  # Merge cells
  merge_v(j = "state") %>%
  merge_v(j = c("appliance", "refClass")) %>%
  # Conditional formatting for both family and count
  color(i = ~ is_correct == TRUE, j = c("family", "n"), color = "darkgreen") %>%
  color(i = ~ is_correct == FALSE, j = c("family", "n"), color = "darkred") %>%
  bg(i = ~ is_correct == TRUE, j = c("family", "n"), bg = "#e6ffe6") %>%
  bg(i = ~ is_correct == FALSE, j = c("family", "n"), bg = "#ffe6e6") %>%
  # Style improvements
  theme_vanilla() %>%
  align(j = c("family", "n"), align = "right") %>%
  border_outer() %>%
  border_inner() %>%
  fontsize(size = 11) %>%
  padding(padding = 4) %>%
  # Add section borders
  hline(i = ~ state != lag(state, default = first(state)), 
        border = officer::fp_border(color = "gray40", width = 2)) %>%
  colformat_double(j = "family", digits = 1) %>%
  set_header_labels(
    state = "State",
    appliance = "Appliance",
    family = "Family",
    refClass = "Reference Class",
    n = "Count",
    is_correct = "Correct Value"
  ) %>%
  autofit()

@@

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(distinct_state_avg = n_distinct(state_avg)) %>%
  ungroup() %>%
  filter(distinct_state_avg > 1) %>%
  group_by(appliance, state_avg, state, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  # Identify outliers (rows where n is 1)
  mutate(is_outlier = n == 1) %>%
  # Reorganize and format
  select(state, refClass, appliance, state_avg, n) %>%
  arrange(state, appliance, refClass, state_avg) %>%
  flextable() %>%
  # Add title
  add_header_lines(values = "State Avg. per utility counts (green=most frequent)") %>%
  bold(part = "header", i = 1) %>%
  # Merge cells
  merge_v(j = "state") %>%
  merge_v(j = c("appliance", "refClass")) %>%
  # Conditional formatting for outliers
  color(i = ~ n == 1, j = c("state_avg", "n"), color = "darkred") %>%
  color(i = ~ n > 1, j = c("state_avg", "n"), color = "darkgreen") %>%
  bg(i = ~ n == 1, j = c("state_avg", "n"), bg = "#ffe6e6") %>%
  bg(i = ~ n > 1, j = c("state_avg", "n"), bg = "#e6ffe6") %>%
  # Style improvements
  theme_vanilla() %>%
  align(j = c("state_avg", "n"), align = "right") %>%
  border_outer() %>%
  border_inner() %>%
  fontsize(size = 11) %>%
  padding(padding = 4) %>%
  # Add section borders
  hline(i = ~ state != lag(state, default = first(state)), 
        border = officer::fp_border(color = "gray40", width = 2)) %>%
  colformat_double(j = "state_avg", digits = 0) %>%
  set_header_labels(
    state = "State",
    appliance = "Appliance",
    state_avg = "State Average",
    refClass = "Reference Class",
    n = "Count"
  ) %>%
  autofit()
```
